### 2023-02-03 GoalDocument.tex
Reformat involved parties

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[UKenglish]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=false, allcolors=blue]{hyperref}

\title{Goal Document: Translating Typing Rules into JastAdd}
\author{Nicholas Boyd Isacsson}

\begin{document}
\maketitle

\section{Involved parties}
\subsubsection*{Student}
  Nicholas Boyd Isacsson \\
  \href{mailto:nicholas@isacsson.se}{nicholas@isacsson.se}

\subsubsection*{Supervisor}
  Christoph Reichenbach \\
  \href{mailto:christoph.reichenbach@cs.lth.se}{christoph.reichenbach@cs.lth.se}

\subsubsection*{Examiner}
  Niklas Fors \\
  \href{mailto:niklas.fors@cs.lth.se}{niklas.fors@cs.lth.se}

\section{Background}

\section{Project description}

\section{Motivation}

\section{Previous work}

\section{Evaluation}

\section{Methodology}

\end{document}