### 2023-02-03 GoalDocument.tex
Add comments about what to cover

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[UKenglish]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=false, allcolors=blue]{hyperref}

\title{Goal Document: Translating Typing Rules into JastAdd}
\author{Nicholas Boyd Isacsson}

\begin{document}
\maketitle

\section{Involved parties}
\subsubsection*{Student}
  Nicholas Boyd Isacsson \\
  \href{mailto:nicholas@isacsson.se}{nicholas@isacsson.se}

\subsubsection*{Supervisor}
  Christoph Reichenbach \\
  \href{mailto:christoph.reichenbach@cs.lth.se}{christoph.reichenbach@cs.lth.se}

\subsubsection*{Examiner}
  Niklas Fors \\
  \href{mailto:niklas.fors@cs.lth.se}{niklas.fors@cs.lth.se}

\section{Background}

\section{Project description}
% deliniate scope, syntax-directed rules, need not be fully self-contained
% eventually break out into one of non syntax-directed rules Hindl-Miller
% or subtyping and adhoc polymorphism

%\section{Motivation} % merge with background
% Why compile to JastAdd?

\section{Previous work}

\section{Evaluation}
%  Case studies, implementation possible by hand or already finished implementations
%  Compare with handwritten implementations:
%    - Spec size
%    - Performance
%  Compare with datalog paper

\section{Methodology}
% Steps:
%   - Input language and parsing
%   - Simply typed lambda calculus
%   - Additional features
%   ...
%   - Evaluation
%   - Finalise the thesis document (collect bullet points throughout and organise them but don't write text)

\end{document}